---
title: "Working with included data sets"
output:
  html_document: 
    fig_height: 6
    fig_width: 10
    theme: cerulean
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      number_sections: true
---

```{r klippy, echo=FALSE, include=TRUE}
# Use this to add a copy-code buttong
klippy::klippy(position = c('top', 'right'), color = 'darkred')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE, cache=FALSE)
```

```{r message=FALSE, warning=FALSE}
library(emln)
library(igraph)
library(bipartite)
library(magrittr)
library(ggplot2)
library(readr)
library(tibble)
library(dplyr)
library(readr)
library(stringr)
library(ggsci)
library(ggbreak)
library(ggpubr)
```

# About the data sets

```{r include=FALSE, echo=FALSE}
paper_figs_theme <- 
  theme_bw()+
  theme(panel.grid = element_blank(),
        panel.border = element_rect(color = "black",fill = NA,size = 1),
        panel.spacing = unit(0.5, "cm", data = NULL),
        axis.text = element_text(size=14, color='black'),
        axis.title = element_text(size=14, color='black'),
        axis.line = element_blank())
paper_figs_theme_no_legend <- 
  paper_figs_theme +
  theme(legend.position = 'none')

## creation of the lists
#read the file path
parent.folder <- "../data/data-raw"
getwd()

sub.folders.short <- list.dirs(path = parent.folder, recursive = F, full.names = F)
sub.folders.long <- list.dirs(path = parent.folder, recursive = F, full.names = T)

#order the sub folders according to the id, using the short sub folders so I can use the id, but I eventually need the long to read the csvs
sub.folders.long <- sub.folders.long[str_order(sub.folders.short, numeric = T)]

## the number of multilayer networks in the database
number_of_networks <- length(sub.folders.long)

id_readout <- data.frame(id = 1:number_of_networks, 
                         network_name = sub.folders.short[str_order(sub.folders.short, numeric = T)], 
                         network_path = sub.folders.long)

#create big lists that each contain the nodes/layers/description/references csv
#NOTE! not all networks have a references csv
all_layers<-list()
all_interactions<-list()
all_nodes<-list()
all_references<-list()
all_descriptions<-list()
all_state_nodes <- list()

##read the relevant csv's in each netowrk to retrieve the information
for(i in 1:number_of_networks) {
    all_layers[[i]]<-read.csv(paste0(sub.folders.long[i],"/layers.csv"))
    all_interactions[[i]]<-read.csv(paste0(sub.folders.long[i],"/interactions.csv"))
    all_nodes[[i]]<-read.csv(paste0(sub.folders.long[i],"/nodes.csv"))
    all_descriptions[[i]]<-read.csv(paste0(sub.folders.long[i],"/description.csv"))
  if (file.exists(paste0(sub.folders.long[i],"/references.csv"))) {
    all_references[[i]]<-read.csv(paste0(sub.folders.long[i],"/references.csv"))
  } else {
    all_references[[i]] <- NA #if the network doesn't have a reference csv (it was not published)
  }
  if (file.exists(paste0(sub.folders.long[i],"/state_nodes.csv"))) {
    all_state_nodes[[i]] <- read.csv(paste0(sub.folders.long[i],"/state_nodes.csv"))
  } else {
    all_state_nodes[[i]] <- NA
  }
}


## lists that contain the type of interactions, nodes, networks and layers
layer_type<-NULL 
node_type<-NULL
interaction_type<-NULL
multilayer_type<-NULL

## lists that contain the number of interactions, nodes, and layers per network
num_of_layers<-c()
num_of_nodes<-c()
num_of_interactions<-c()

## lists that contain the unique attributes of interactions, nodes and layers
layer_attribute<-NULL
attributes_interactions<-NULL
attributes_nodes<-NULL

##create the lists using the files that contain the types, number, ettributes etc...##
# Issue 85 - Network 39 does not have an appropriate $description format #
# Issue 50 - Network 39 does not have an appropriate $description format #
for(i in 1:number_of_networks){ 
    layer_type<-rbind(layer_type,tibble(network_id=i,filter(all_layers[[i]],attribute=="type")))
    node_type<-rbind(node_type,tibble(network_id=i,filter(all_nodes[[i]],attribute =="type")))
    interaction_type<-rbind(interaction_type,tibble(network_id=i,filter(all_interactions[[i]],attribute=="type")))
    num_of_layers<-append(num_of_layers,max(all_layers[[i]]$layer))
    num_of_nodes<-append(num_of_nodes,max(all_nodes[[i]]$node_id))
    num_of_interactions<-append(num_of_interactions,max(all_interactions[[i]]$interaction_id))
    layer_attribute<-rbind(layer_attribute,tibble(network_id=i,att=all_layers[[i]][["attribute"]]))
    attributes_interactions<-rbind(attributes_interactions,tibble(network_id=i,att=all_interactions[[i]][["attribute"]]))
    attributes_nodes<-rbind(attributes_nodes,tibble(network_id=i,att=all_nodes[[i]][["attribute"]]))
    #multilayer_type<-rbind(network_type,tibble(network_id=i,filter(all_descriptions[[i]],attribute=="type")))
    
  }
```

```{r network_information, echo=FALSE}
###multilayer network types###
multi <- data.frame()
for (i in 1:length(all_descriptions)) {
  # Issue 85 - Network 39 does not have an appropriate $description format #
  #multi <- rbind(multi, data.frame(network_id = i,
  #               multi_type = all_descriptions[[i]][all_descriptions[[i]]$attribute == 'multilayer_network_type',]$value))
  subset_df <- all_descriptions[[i]][all_descriptions[[i]]$attribute == 'multilayer_network_type', ]
  
  if (nrow(subset_df) == 0) {
    row <- data.frame(network_id = i, multi_type = "Missed")
  } else {
    row <- data.frame(network_id = i, multi_type = subset_df$value)
  }
  
  multi <- rbind(multi, row)
}


#count the number of types present in all the networks
counted_multi_types<-count(x = multi,multi_type)
colnames(counted_multi_types) = c('multi_type','count')

##create a dataframe that contains the proportion of each multi type for visualization
counted_multi_types <- counted_multi_types %>% 
  arrange(desc(multi_type)) %>%
  mutate(prop = count / sum(counted_multi_types$count) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )


###ecological layer types###
eco <- data.frame()
for (i in 1:length(all_descriptions)) {
  #eco <- rbind(eco, data.frame(network_id = i,
  #               eco_type = all_descriptions[[i]][all_descriptions[[i]]$attribute == 'ecological_network_type',]$value))
   subset_df <- all_descriptions[[i]][all_descriptions[[i]]$attribute == 'ecological_network_type', ]
  
  if (nrow(subset_df) == 0) {
    row <- data.frame(network_id = i, eco_type = "Missed")
  } else {
    row <- data.frame(network_id = i, eco_type = subset_df$value)
  }
  
  eco <- rbind(eco, row)
}


#count the number of types present in all the networks
counted_eco_types<-count(x = eco,eco_type)
colnames(counted_eco_types) = c('eco_type','count')

##create a dataframe that contains the proportion of each multi type for visualization
counted_eco_types <- counted_eco_types %>% 
  arrange(desc(eco_type)) %>%
  mutate(prop = count / sum(counted_eco_types$count) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )


#combination of layer type and network type
multi_eco <- merge(multi,eco,by = 'network_id')

multi_eco <- count(multi_eco, multi_type, eco_type)

multi_eco$multi_type <- factor(multi_eco$multi_type , levels=c("Spatial","Temporal","Environment", "Multiplex","Perturbation"))
```

The package includes `r number_of_networks` data sets of multilayer networks, collected from different sources. This collection of data is highly useful for comparative analysis and for practicing workflows. How data were collected and organized is described in the paper.

The data includes `r nrow(counted_multi_types)` types of multilayer networks:

```{r, echo=FALSE}
counted_multi_types %>% arrange(desc(count)) %>% select(multi_type, count)
```

These are divided into `r nrow(counted_eco_types)` types of ecological interaction networks:. 

```{r, echo=FALSE}
counted_eco_types %>% arrange(desc(count)) %>% select(eco_type, count)
```

Here is a visual summary:

```{r, echo=FALSE, fig.height=8}
ggplot(multi_eco, aes(x=multi_type,y=n, fill=eco_type)) + 
  geom_bar(stat = 'identity', color='black') + xlab('Multilayer Network Type') + paper_figs_theme +
  theme(axis.text.x = element_text(angle = 75, vjust = 0.5, size = 12), legend.position = 'bottom') + 
  ylab('Count')  + scale_fill_nejm(name = 'Ecological Network Type')
```

The data is distributed globally:

```{r Map_Visualization, echo=FALSE}
latitudes<-NULL
longitudes<-NULL
coord <- NULL

for(i in 1:number_of_networks){
  if ('latitude' %in% all_layers[[i]]$attribute) {
    coord <- rbind(coord, data.frame(network_id = i,
                                     latitude = as.numeric(unlist(strsplit(filter(all_layers[[i]],
                                                                       attribute == 'latitude')$value,split = ','))),
                                     longitude = as.numeric(unlist(strsplit(filter(all_layers[[i]],
                                                                        attribute == 'longitude')$value, split =',')))
    ))
    latitudes<-c(latitudes,all_layers[[i]] %>% filter(attribute=="latitude") %>% unique)
    longitudes<-c(longitudes,all_layers[[i]] %>% filter(attribute=="longitude") %>% unique) 
  }
}

coord <- coord %>%
  group_by(latitude, longitude) %>%
  distinct() %>% 
  na.omit()


#there is a problem with the coordinates of network 39
coord <- coord[!coord$network_id==39,]

#add attributes to the map
coord <- merge(coord,multi[,c('network_id','multi_type')], by = 'network_id')

worldmap <- map_data("world")
ggplot(worldmap) +
  geom_map(
    data = worldmap, map = worldmap,
    aes(long, lat, map_id = region),
    color = "white", fill = "lightgray", size = 0.1
  ) +
  geom_point(
    data = coord,
    aes(longitude, latitude, color = multi_type),
    alpha = 0.6) + expand_limits(x = worldmap$long, y = worldmap$lat) +
  theme_void() + theme(title = NULL, legend.position = 'bottom')  + scale_color_nejm() +labs(color='Multilayer type')
```

# Browsing the data

You can browse through the data sets using `view_emln()` in RStudio. The function will open a user interface in the viewer panel. To search for a network with a certain set of parameters, you can use `search_emln()`. Here are a few examples.

```{r}
search_1 <- search_emln(ecological_network_type = 'Host-parasite')
names(search_1)
# Show some of the columns
search_1 %>% select("network_id","network_name","ecological_network_type","multilayer_network_type","layer_num")

search_2 <- search_emln(ecological_network_type = 'Pollination', layer_number_minimum = 15)
search_2 %>% select("network_id","network_name","ecological_network_type","multilayer_network_type","layer_num")

# Networks with state node attributes
search_emln(state_nodes = TRUE)

# Networks with interlayer links
search_emln(interlayer = TRUE)
```

# Loading a network

We load the spatial multilayer network (network_id = 38) that comes with the package. This is a spatial seed-Dispersal network (collected from Web of Life). The network has five layers and 22 physical nodes, with a total of 84 interactions that are all intralayer, undirected, and weighted. There are no interlayer interactions recorded for this network.

```{r}
d38 <- load_emln(38)
names(d38)
d38$layers
head(d38$extended)
```

<!-- Here is another example: -->

<!-- ```{r} -->
<!-- search_emln(ecological_network_type = 'Pollination', directed = TRUE) -->
<!-- d47 <- load_emln(47) -->
<!-- d47 -->
<!-- ``` -->

