---
title: "Handling monolayer network data"
output:
  html_document: 
    fig_height: 4
    fig_width: 6
    theme: cerulean
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      number_sections: true
---


```{r klippy, echo=FALSE, include=TRUE}
# Use this to add a copy-code buttong
klippy::klippy(position = c('top', 'right'), color = 'darkred')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE, cache=FALSE)
```

```{r message=FALSE, warning=FALSE}
library(emln)
library(igraph)
library(tidyverse)
library(bipartite)
library(magrittr)
```


We provide examples with simulated toy networks and with real data to create an object of class `monolayer`. See help (`?monolayer`) to learn more about this class 

# Unipartite 

We start with a toy example. We generate a 3x3 matrix with random weighted interactions.

```{r monolayer_uni_1}
n <- 3
adjacency_matrix <- matrix(round(runif(n*n,1,5),0), ncol = n, dimnames = list(paste0("A", 1:n), paste0("A", 1:n)))

# Convert a matrix to a monolayer object
matrix_to_list_unipartite(adjacency_matrix, directed = TRUE)
```

Now let's see a toy example for converting a link list.

```{r monolayer_uni_2}
# Generate a data frame with random links
link_list_unipartite <- data.frame(from=c("A1", "A1", "A1", "A2", "A2", "A2", "A3", "A3","A3"), to=c("A1", "A2", "A3", "A1", "A2", "A3", "A1", "A2","A3"), weight=round(runif(n*n,1,5),0))

list_to_matrix(link_list_unipartite, directed = TRUE, bipartite = FALSE)
```

To illustrate working with real data we will use a binary directed marine food web from [Jacob, Ute, Aaron Thierry, Ulrich Brose, Wolf E. Arntz, Sofia Berg, Thomas Brey, Ingo Fetzer, et al. 2011. “The Role of Body Size in Complex Food Webs: A Cold Case.” In Advances in Ecological Research, edited by Andrea Belgrano, 45:181–223. Academic Press.](https://www.sciencedirect.com/science/article/pii/B9780123864758000058) Kongsfjorden is a glacial fjord on the northwest corner of the Svalbard archipelago. It is a 30 km open fjord with no marked sill at the entrance, and with a maximum depth exceeding 300 m. The network consists of 262 species with 1,544 feeding interactions. **Note the warning messages**; these are OK for this data set.

**These data are organized as a link list, so it is possible to include node attributes.**


```{r monolayer_uni_data_1}
# The data is included in the infomapecology package
data("kongsfjorden_links", package = 'infomapecology')
data("kongsfjorden_nodes", package = 'infomapecology')
nodes <- kongsfjorden_nodes %>%
  select(node_name=Species, node_id_original=NodeID, everything())

interactions <- kongsfjorden_links %>%
  select(from=consumer, to=resource) %>%
  mutate_if(is.factor, as.character) %>%
  mutate(weight=1)

monolayer_unipartite <- list_to_matrix(x=interactions, directed = T, bipartite = F, node_metadata = nodes)

head(monolayer_unipartite$nodes)

head(monolayer_unipartite$edge_list)
```

Now let's take the matrix we obtained as an example, and create a monolayer object from a matrix.

```{r monolayer_uni_data_2}
kongsfjorden_matrix <- monolayer_unipartite$mat
monolayer_unipartite <- matrix_to_list_unipartite(x = kongsfjorden_matrix, directed = T)
```



# Bipartite

We start with a toy example. We generate a random incidence matrix with dimensions P X A. Bipartite networks have two sets of nodes with no overlap between the sets.

```{r monolayer_bip_1}
# Case "Bipartite networks"
# 
P <- 4 # 4 rows
A <- 3 # 3 columns
incidence_matrix <- matrix(round(runif(P*A,1,5),0), ncol = A, dimnames = list(paste0("P", 1:P), paste0("A", 1:A)))

# Convert an matrix to an edge list for bipartite network
matrix_to_list_bipartite(incidence_matrix, group_names = c('A_set', 'P_set'))$edge_list
```

To demonstrate real data we use the incidence matrix olesen2002flores included in the bipartite package.

```{r monolayer_bip_2}
# create a monolayer object from a bipartite matrix
monolayer_bipartite <- matrix_to_list_bipartite(x = olesen2002flores, group_names = c('Pollinator', 'Plant'))

# What is included in the monolayer class object
names(monolayer_bipartite)

# The nodes in the network
monolayer_bipartite$nodes

# The link (edge) list
monolayer_bipartite$edge_list

# An igraph object is also created
monolayer_bipartite$igraph_object
```

We can also convert a link list to a monolayer object.

```{r monolayer_bip_3}
ll <- monolayer_bipartite$edge_list
monolayer_bipartite_2 <- list_to_matrix(ll, bipartite = T, group_names = c('Pollinator', 'Plant'))
```


# Function `create_monolayer_network`

This function is a wrapper for the `matrix_to_list_unipartite`, `matrix_to_list_bipartite`, and `list_to_matrix` functions. It is convenient because it automatically selects the input type.

```{r create_monolayer_network}
#Input is a link list
monolayer_unipartite <- create_monolayer_network(x=interactions, directed = T, bipartite = F, node_metadata = nodes)

# Input is a matrix
bipartite_monolayer <- create_monolayer_network(x = olesen2002flores, group_names = c('Pollinator', 'Plant'), bipartite = T)
```

