---
title: "Visualizing multilayer networks"
output:
  html_document:
    fig_height: 4
    fig_width: 6
    theme: cerulean
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      number_sections: true
---


```{r klippy, echo=FALSE, include=TRUE}
# Use this to add a copy-code button
klippy::klippy(position = c('top', 'right'), color = 'darkred')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE, cache=FALSE)
```

```{r message=FALSE, warning=FALSE}
library(emln)
```

The `emln` package ships with an interactive **Multilayer Network Visualizer** — a browser-based web app for exploring multilayer networks in multiple views. You can launch it directly from R with `plot_multilayer()`, which starts a local server and opens the app in your browser. Alternatively, export your network to JSON or CSV and upload it to the [standalone version of the app](https://mira.ecomplab.com/).

> **Note:** The Multilayer Network Visualizer is currently in **beta**. Feedback and bug reports are welcome at the [emln GitHub issues page](https://github.com/Ecological-Complexity-Lab/emln/issues).

# Load a dataset

We will use network 60 from the built-in emln dataset collection — a bipartite multilayer network. Load it with `load_emln()` and inspect its reference.

```{r load_data}
net <- load_emln(60)
net$references
```

The result is a `multilayer` object. Its three core tables are the physical nodes, the layers, and the extended edge list (one row per interaction, identified by the layer and node on each end).

```{r inspect_data}
# Physical nodes (first few rows)
head(net$nodes)

# Layers
net$layers

# Extended edge list (first few rows)
head(net$extended)
```

Note that interlayer edges (where `layer_from != layer_to`) encode connections between layers, while intralayer edges (`layer_from == layer_to`) encode connections within a layer.

# Interactive visualization with `plot_multilayer()`

`plot_multilayer()` converts the multilayer object to JSON, starts a local HTTP server using `httpuv`, and opens the visualizer in your browser. The network data is held in memory — nothing is written to disk. The function returns the server handle, which you use to stop the server when you are done.

**Important:** `bipartite` is not auto-detected; you must set it explicitly. The `directed` argument defaults to `NULL` and is auto-detected from edge symmetry.

```{r plot_multilayer, eval=FALSE}
# Open the visualizer in the browser
srv <- plot_multilayer(net, bipartite = TRUE, browser = 'Chrome')

# When done, stop the local server
httpuv::stopServer(srv)
```

The app opens at `http://localhost:8080?autoload=true` (the port increments automatically if 8080 is busy). It provides four main views:

- **Network view** — a 3D stacked-layer canvas. Rotate, pan, drag nodes, zoom, and hover to inspect individual nodes and links.
- **Map view** — layers positioned on an interactive world map when latitude and longitude coordinates are present in the layers table.
- **Layer view** — a meta-graph where each layer is a bubble; click a bubble to open a side-by-side comparison of any two layers.
- **Dashboard view** — analytics panels including degree distributions, link weight distributions, a presence/absence matrix across layers, a Jaccard similarity heatmap between layers, and more.

# Export to JSON with `multilayer_to_json()`

Use `multilayer_to_json()` to save your network as a JSON file compatible with the visualizer's **Import** button. This is useful when you want to share the network, reload it later without re-running R, or upload it to the standalone app.

```{r to_json, eval=FALSE}
multilayer_to_json(net, file = "net60.json", bipartite = TRUE)
```

If you omit `file`, the function returns the JSON string — useful for passing it to another function or inspecting the output programmatically.

```{r to_json_string, eval=FALSE}
json_str <- multilayer_to_json(net, bipartite = TRUE)
cat(substr(json_str, 1, 500))   # preview the first 500 characters
```

The JSON object contains four arrays that the visualizer expects:

| Array | Contents |
|-------|----------|
| `nodes` | Physical nodes: `node_id`, `node_name`, and any extra node attributes |
| `layers` | Layer metadata: `layer_id`, `layer_name`, and any extra layer attributes. For bipartite networks a `bipartite: true` flag is added. |
| `extended` | Extended edge list: `layer_from`, `node_from`, `layer_to`, `node_to`, `weight`, and any extra link attributes |
| `state_nodes` | State node map: `layer_id`, `node_id`, `layer_name`, `node_name` |

A top-level `directed` flag is set automatically based on edge symmetry (unless you override it with the `directed` argument).

# Export to CSV with `multilayer_to_csv()`

If you prefer a tabular format, `multilayer_to_csv()` writes three or four CSV files to a directory of your choice, following the schema accepted by the visualizer's CSV importer.

```{r to_csv, eval=FALSE}
multilayer_to_csv(net, dir = "out/", prefix = "net60", bipartite = TRUE)
```

The files written are:

| File | Contents |
|------|----------|
| `net60_edges.csv` | Extended edge list: `layer_from`, `node_from`, `layer_to`, `node_to`, `weight`, plus any extra link attributes |
| `net60_layers.csv` | Layer metadata: `layer_id`, `layer_name`, plus any extra layer attributes (and a `bipartite` column for bipartite networks) |
| `net60_nodes.csv` | Physical node attributes: `node_name`, `node_type`, plus any extras that are constant across layers |
| `net60_state_nodes.csv` | Per-(layer, node) attributes such as abundance or module membership — **written only when such attributes exist** in the multilayer object |

The `directed` flag is not stored in the CSV files. You specify it in the visualizer's CSV import dialog at load time.

To import the CSV files, open the visualizer, click **Import CSV**, select your three (or four) files, and confirm the `directed` setting.

# Using the standalone visualizer

The visualizer is also available as a standalone web app at [https://mira.ecomplab.com/](https://mira.ecomplab.com/) — no R session required. Export your network to JSON or CSV from R, then upload it there to share or explore it in any browser.
