---
title: "Visualizing multilayer networks"
output:
  html_document:
    fig_height: 4
    fig_width: 6
    theme: cerulean
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      number_sections: true
---


```{r klippy, echo=FALSE, include=TRUE}
# Use this to add a copy-code button
klippy::klippy(position = c('top', 'right'), color = 'darkred')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE, cache=FALSE)
```

```{r message=FALSE, warning=FALSE}
library(emln)
```

<img src="MiRA_logo.svg" alt="MiRA logo" style="float:right; width:100px; margin-left:20px; margin-right:5px; margin-bottom:10px;" />

The `emln` package ships with MiRA, an interactive **Multilayer Network Visualizer** — a browser-based web app for exploring multilayer networks in multiple views. You can launch it directly from R with `plot_multilayer()`, which starts a local server and opens the app in your browser. Alternatively, export your network to JSON or CSV and upload it to the [standalone version of the app](https://mira.ecomplab.com/).

> **Note:** Feedback and bug reports for MiRA are welcome at the [MiRA GitHub issues page](https://github.com/Ecological-Complexity-Lab/MiRA/issues).

If you use MiRA in published research, please cite: Nehorai S, Bloch Y and Pilosof S. *Interactively visualizing biological multilayer networks using MiRA.* (forthcoming)


# Interactive visualization

`plot_multilayer()` converts the multilayer object to JSON, starts a local HTTP server using `httpuv`, and opens MiRA in your browser. The network data is held in memory — nothing is written to disk. The function returns the server handle, which you use to stop the server when you are done.

**Important:** `bipartite` is not auto-detected; you must set it explicitly. The `directed` argument defaults to `NULL` and is auto-detected from edge symmetry.

Two additional optional arguments are worth knowing:

- `setA_type` — for bipartite networks, specifies which `node_type` value is displayed as Set A (the top row). By ecological convention this is the higher trophic level (e.g. `"pollinator"`, `"parasite"`). If omitted, MiRA falls back to alphabetical ordering.
- `directed_interlayer` — controls whether interlayer links are directed independently from intralayer links. Useful for temporal multilayer networks where interlayer links represent time progression. If `NULL` (default), it inherits the value of `directed`.

```{r plot_multilayer, eval=FALSE}
# Load data
net <- load_emln(65)

# Open MiRA in the browser
srv <- plot_multilayer(net, bipartite = TRUE, setA_type = "pollinator", browser = 'Chrome')

# When done, stop the local server
httpuv::stopServer(srv)
```

The app opens at `http://localhost:8080?autoload=true` (the port increments automatically if 8080 is busy). MiRA provides seven visualization modes:

- **Network mode** — a 3D stacked-layer canvas. Rotate, pan, drag nodes, zoom, and hover to inspect individual nodes and links.
- **Map mode** — layers positioned on an interactive world map when latitude and longitude coordinates are present in the layers table.
- **Grid view** — a small-multiples layout that shows all layers simultaneously as individual panels with a shared colour scheme and selection state; useful for quick visual comparison across many layers.
- **Layer view** — a meta-graph where each layer is a force-directed bubble with a micro-graph preview; click a bubble to open a side-by-side comparison of any two layers.
- **Meta-network mode** — aggregates all layers into a single-layer view that highlights cross-layer connectivity patterns.
- **Dashboard mode** — analytics panels including degree distributions, link weight distributions, a presence/absence matrix across layers, a Jaccard similarity heatmap between layers, and more.
- **Data mode** — a tabular inspector for nodes, links, and layer attributes; supports filtering, subsetting, and CSV export directly from the browser.

# Export to JSON

Use `multilayer_to_json()` to save your network as a JSON file compatible with the visualizer's **Import** button. This is useful when you want to share the network, reload it later without re-running R, or upload it to the standalone app.

```{r to_json, eval=FALSE}
multilayer_to_json(net, file = "net60.json", bipartite = TRUE)
```

If you omit `file`, the function returns the JSON string — useful for passing it to another function or inspecting the output programmatically.

```{r to_json_string, eval=FALSE}
json_str <- multilayer_to_json(net, bipartite = TRUE)
cat(substr(json_str, 1, 500))   # preview the first 500 characters
```

The JSON object contains four arrays that the visualizer expects:

| Array | Contents |
|-------|----------|
| `nodes` | Physical nodes: `node_id`, `node_name`, and any extra node attributes |
| `layers` | Layer metadata: `layer_id`, `layer_name`, and any extra layer attributes. For bipartite networks a `bipartite: true` flag is added; if `setA_type` is provided it is also written per layer. |
| `extended` | Extended edge list: `layer_from`, `node_from`, `layer_to`, `node_to`, `weight`, and any extra link attributes |
| `state_nodes` | State node map: `layer_id`, `node_id`, `layer_name`, `node_name`, plus any extra per-(layer, node) attributes (e.g. abundance, module membership) |

Two top-level flags control edge directionality: `directed` (intralayer) is set automatically from edge symmetry unless overridden; `directed_interlayer` is written only when set explicitly, and defaults to inheriting from `directed`.

# Export to CSV

If you prefer a tabular format, `multilayer_to_csv()` writes three or four CSV files to a directory of your choice, following the schema accepted by the visualizer's CSV importer.

```{r to_csv, eval=FALSE}
multilayer_to_csv(net, dir = "out/", prefix = "net60", bipartite = TRUE)
```

The files written are:

| File | Contents |
|------|----------|
| `net60_edges.csv` | Extended edge list: `layer_from`, `node_from`, `layer_to`, `node_to`, `weight`, plus any extra link attributes |
| `net60_layers.csv` | Layer metadata: `layer_id`, `layer_name`, plus any extra layer attributes (and `bipartite` and `setA_type` columns for bipartite networks) |
| `net60_nodes.csv` | Physical node attributes: `node_name`, `node_type`, plus any extras that are constant across layers |
| `net60_state_nodes.csv` | Per-(layer, node) attributes such as abundance or module membership — **written only when such attributes exist** in the multilayer object |

The `directed` flag is not stored in the CSV files. You specify it in the visualizer's CSV import dialog at load time.

To import the CSV files, open the visualizer, click **Import CSV**, select your three (or four) files, and confirm the `directed` setting.

# Using the standalone visualizer

The visualizer is also available as a standalone web app at [https://mira.ecomplab.com/](https://mira.ecomplab.com/) — no R session required. Export your network to JSON or CSV from R, then upload it there to share or explore it in any browser.
